Phenotype pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues

### 2022-08-19 ms/ms01_main.tex
ENH added crm analysis

eQTLs

@@

\subsection*{Which regulatory variants are the most pleiotropic?}

@@

\subsection*{What is the function of pleiotropic variants?}

@@

\subsection*{Analysis of pleiotropic regions}

@@

@@

Then I hypothesized that pleiotropic variants modulate more egenes.
%
I found that pleiotropic variants have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
%
This observation could be explained by my previous observation that pleiotropic variants have more often an effect on the splicing and 3'UTR regions (Figure \ref{fig:vep_consequence}b,c).

@@

% Confirm eQTL studies
In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 125 eQTL studies.
%
This pipeline and the results is useful for three aspects.
%
First these results allow to investigate the molecular characteristics of pleiotropic variants with a gene regulatory effect.
%
Second, the colocalized loci provide predictions of causal variants in given loci of GWAS.
%
% TODO how many
Third, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of many GWAs traits.

In this work, I have explore variants with a role in gene regulation that are highly pleiotropic.
%
For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.

For a posterior probability $\ge$ 0.8 that the GWAS variants and the eQTLs are common, we found TODO variants.

>>>>

@@

\subsection*{Pleiotropic variants colocalized with eQTLs}

\subsubsection*{Which variants are the most pleiotropic?}

@@

\subsubsection*{What is the function of pleiotropic variants?}

@@

\subsection*{Pleiotropic regions}

@@

%TODO cumulated covering of genome by pleitopic

@@

%TODO egene per variant-etissue
Then I hypothesized that pleiotropic variants modulate more egenes.
%
I found that pleiotropic variants have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
%
This observation could be explained by my previous observation that pleiotropic variants have more often an effect on the splicing and 3'UTR regions (Figure \ref{fig:vep_consequence}b,c).

@@

%TODO idea that gene expression is causal to disease

% Confirm eQTL studies
In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 125 eQTL studies.
%
For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
%
The results of this pipeline provides three types of insights.
%
First these results allow to investigate the molecular characteristics of pleiotropic variants with a gene regulatory effect.
%
Second, the colocalized loci provide predictions of causal variants in given loci of GWAS.
%
% TODO how many
Third, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of many GWAs traits.

% First insight: pleiotropy %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For a posterior probability $\ge$ 0.8 that the GWAS variants and the eQTLs are common, we found TODO variants.
%
Our analysis suggests that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.

These variants are under control of more transcription factors

### 2022-08-19 ms/ms02_floats.tex
ENH added crm analysis

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/missense_variant.png}
%
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/splice_region_variant.png}
%
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/3_prime_UTR_variant.png}
%
\end{subfigure}
%
\caption{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: \textbf{a}, Missense variant, \textbf{b}, splice region variant and \textbf{c}, 3' UTR variant.} \label{fig:vep_consequence}
%
\end{figure*}

@@

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_y_egene.py/plt.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_egene_y_etissue.py/plt.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_egene_etissue_y_gwas.py/plt.png}
\end{subfigure}
%
\caption{Count of egenes per variant (\textbf{a}), etissues per variant-egene (\textbf{b}) and GWAS categories per variant-egene-etissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
%
\end{figure*}

@@

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
%
\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remaptf.py/bxplt_remaptf_per_rsid_flank_0.png}
\end{subfigure}
%
\caption{Binding of transcription factors in the region around pleiotropic variants (Radius 50 bp) as a function of the GWAS category count.} \label{fig:freq_tf_per_variant}
%
\end{figure*}

>>>>

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/missense_variant.png}
%
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/splice_region_variant.png}
%
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/3_prime_UTR_variant.png}
%
\end{subfigure}
%
\caption{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: \textbf{a}, Missense variant, \textbf{b}, splice region variant and \textbf{c}, 3' UTR variant.} \label{fig:vep_consequence}
%
\end{figure*}

@@

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_etissue_y_egene.py/plt.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_egene_y_etissue.py/plt.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_egene_etissue_y_gwas.py/plt.png}
\end{subfigure}
%
\caption{Count of egenes per variant-tissue (\textbf{a}), etissues per variant-egene (\textbf{b}) and GWAS categories per variant-egene-etissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
%
\end{figure*}

@@

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remaptf.py/bxplt_remaptf_per_rsid_flank_0.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_remap_crm_gwas_categories.py/remap_crm_fisher.png}
\end{subfigure}
%
\caption{Analysis of transcription factor binding. \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic variants (Radius 50 bp) as a function of the GWAS category count. \textbf{b}, Odds ratio of cis regulatory module-annotated vs non-annotated variants.} \label{fig:freq_tf_per_variant}
%
\end{figure*}
